Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-10 ms.tex
figure show now reflect new Ia radius prescription
runaway and nscs.

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm inf}$).  However, for low heating the stagnation
radius diverges above the analytic estimate, approaching the stellar
break radius $r_b \gg r_{\rm inf}$.  This divergence occurs
approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which coincides
with the Bernoulli parameter of the gas decreasing below unity near
the stagnation radius (Appendix \ref{app:be}).  This divergence of the
stagnation radius to large radii occurs at a higher value of $\zeta$
in core galaxies ($\Gamma = 0.8$) with large break radii, explaining
the behavior of the two core galaxies shown in Fig.~\ref{fig:stag} as
vertically aligned orange triangles.  These solutions correspond to
the same black hole mass ($\Mbh = 10^8 \Msun$) and heating rate ($v_w
= 600$ km s$^{-1}$) but assume different break radii of $r_{b} =$ 25,
50 pc (from top to bottom).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} Stagnation radius $r_{s}$ in units of the
    sphere of influence radius $r_{\rm inf}$ (eq.~[\ref{eq:rsoi}]) for
    galaxies in our sample as a function of the stellar wind heating
    parameter $\zeta \equiv \sqrt{1+(v_w/\sigma_0)^{2}}$.  Green,
    orange, and blue symbols correspond to different values of $v_{w}
    =$ 300, 600, and 1200 km s$^{-1}$, respectively.  Squares
    correspond to cusp galaxies ($\Gamma = 0.8$), while triangles
    correspond to cores ($\Gamma = 0.1$). Green circles correspond to
    cusp solutions which would be thermally unstable. The black curves
    correspond to the analytic prediction from equation
    (\ref{eq:rs2main}), with thick solid and dot-dashed curves
    calculated for parameters $(\Gamma=0.8, \densSlope\simeq 1)$ and
    $(\Gamma=0.1,\densSlope\simeq0.6)$, respectively. The thin black
    solid line corresponds to the simplified analytic result for $\rs$
    from equation~\eqref{eq:stag_simple} (recall that $\rinf \simeq 14
    \Mbheight^{0.6} {\rm pc}$).}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  Thin solid and dot-dashed curves
    correspond to our analytic estimates of $\dot{M}/\dot{M}_{\rm
      edd}$ (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively. Equation ~[\ref{eq:eddr_analytic}] is derived from
    the approximate analytic result for the stagnation radius given by
    equation~\eqref{eq:stag_simple}. If we instead use the more
    accurate result from equation~\eqref{eq:rs2main} we obtain the
    thick curves.}
\end{figure}

@@

% \begin{figure}
% \includegraphics[width=\columnwidth]{eta.pdf}
% \includegraphics[width=\columnwidth]{rs_rIa.pdf}
% \caption{\label{fig:eta} {\it Top panel:} Parameter $\eta$
%   characterizing stellar mass loss rate (eq.~[\ref{eq:q}]) as a
%   function of black hole mass $\Mbh$, calculated for average star
%   formation histories from \citet{MosterNaab+:2013a} (see
%   Appendix~\ref{app:windheat} for details).  {\it Bottom panel:} Ratio
%   of Ia radius to stagnation radius as a function of $\Mbh$.  The Ia
%   radius calculated from equation~\eqref{eq:rIa} based on convolving
%   the average star formation histories with the SN Ia DTD.  The
%   stagnation radius is calculated from \eqref{eq:stag_simple} using
%   the total heating rate for each $\Mbh$ shown in the top panel. }
% \end{figure}

@@

We conclude by drawing attention to the following limitations of our model.  First, by neglecting any large scale inflow of gas to the nucleus and assuming the only source of gas feeding the black hole is stellar wind mass loss from the local stellar population, our estimates represent a lower bound on the gas density of the CNM.  Our model therefore cannot account for AGN.

@@

On the other hand, \citealt{LauerFaber+:2005a} find that $\sim
      60\%$ of cusp galaxies and $\sim29\%$ of core galaxies have
      (generally unresolved) emission in excess of the inward
      extrapolation of the Nuker law.  They note that these nuclei are bluer
      than the background galaxy starlight, suggesting the presence of
      younger stars towards galactic center {\bf AG mass
        threshold for this effect}.  Mass injection to the nucleus could in some cases therefore be dominated by a nuclear star cluster with its own break radius on parsec scales, i.e. much smaller than our fiducial value of $r_b \sim 100$ pc.  In the case of high heating, a smaller break radius does not appreciably change the stagnation radius, resulting in a similar SMBH accretion rate.  For break radii smaller than the nominal stagnation radius, the presence of a smaller break radius would likely decrease the stagnation radius and accretion rate.  
%      be for large break radii, the stagnation radius would move
 %     inwards somewhat. However note the for high heating rate ($v_w
  %    \gsim 500 $ km/s and small black hole masses $\Mbh \lsim 10^7
   %   \Msun$), we predict stagnation radii smaller than a parsec.

>>>>

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm inf}$).  However, for low heating the stagnation
radius diverges above the analytic estimate, approaching the stellar
break radius $r_b \gg r_{\rm inf}$.  This divergence occurs
approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which coincides
with the Bernoulli parameter of the gas decreasing below unity near
the stagnation radius (Appendix \ref{app:be}).  Thus, for small values
of the heating rate (small $\zeta$) the location of the stagnation
radius will vary strongly with the break radius. This explains the
behavior of the two vertically aligned green circles. These are two cusp
($\Gamma=0.8$) galaxies with the $\Mbh=10^7 \Msun$ and $v_w=300$ km/s
but with different break radii ($\rb=100, 400$ pc from top to
bottom). This divergence of the stagnation radius to large radii
occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.1$),
explaining the behavior of the two core galaxies shown in
Fig.~\ref{fig:stag} as vertically aligned orange triangles.  These
solutions correspond to the same black hole mass ($\Mbh = 10^8 \Msun$)
and heating rate ($v_w = 600$ km s$^{-1}$) but assume different break
radii of $r_{b} =$ 25, 50 pc (from top to bottom).

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} Stagnation radius $r_{s}$ in units of the
    sphere of influence radius $r_{\rm inf}$ (eq.~[\ref{eq:rsoi}]) for
    galaxies in our sample as a function of the stellar wind heating
    parameter $\zeta \equiv \sqrt{1+(v_w/\sigma_0)^{2}}$.  Green,
    orange, and blue symbols correspond to different values of $v_{w}
    =$ 300, 600, and 1200 km s$^{-1}$, respectively.  Squares
    correspond to cusp galaxies ($\Gamma = 0.8$), while triangles
    correspond to cores ($\Gamma = 0.1$). Green circles correspond to
    cusp solutions which would be thermally unstable. The green circle
    with an upwards arrow corresponds to a lower bound--this is a
    model for where we were not able to obtain a converged value for
    the stagnation radius as it moves outwards when the outer boundary
    is moved outwards. The black curves correspond to the analytic
    prediction from equation (\ref{eq:rs2main}), with thick solid and
    dot-dashed curves calculated for parameters $(\Gamma=0.8,
    \densSlope\simeq 1)$ and $(\Gamma=0.1,\densSlope\simeq0.6)$,
    respectively. The thin black solid line corresponds to the
    simplified analytic result for $\rs$ from
    equation~\eqref{eq:stag_simple} (recall that $\rinf \simeq 14
    \Mbheight^{0.6} {\rm pc}$).}
\end{figure}

\begin{figure}
  \includegraphics[width=\columnwidth]{mdot_mass.eps}
  \caption{\label{fig:mdot_mass} Accretion rate $\dot{M}/\dot{M}_{\rm
      edd}$ versus SMBH mass for galaxies in our sample, calculated
    for different values of the wind heating parameter $\vwO =$ 300 km
    s$^{-1}$ ({\it green}), 600 km s$^{-1}$ ({\it orange}), and 1200
    km s$^{-1}$ ({\it blue}).  Squares correspond to cusp galaxies
    ($\Gamma=0.8$), while triangles correspond to cores
    ($\Gamma$=0.1). Green circles correspond to cusp solutions which
    would be thermally unstable.  The green circle with an upwards
    arrow corresponds to a lower bound--this is a model for where we
    were not able to obtain a converged value for the stagnation
    radius as it moves outwards when the outer boundary is moved
    outwards.  Thin solid and dot-dashed curves correspond to our
    analytic estimates of $\dot{M}/\dot{M}_{\rm edd}$
    (eq.~[\ref{eq:eddr_analytic}]) for cusp and core galaxies,
    respectively. Equation ~[\ref{eq:eddr_analytic}] is derived from
    the approximate analytic result for the stagnation radius given by
    equation~\eqref{eq:stag_simple}. If we instead use the more
    accurate result from equation~\eqref{eq:rs2main} we obtain the
    thick curves.}
\end{figure}

@@

@@

We conclude by drawing attention to the following limitations of our
model.  First, by neglecting any large scale inflow of gas to the
nucleus and assuming the only source of gas feeding the black hole is
stellar wind mass loss from the local stellar population, our
estimates represent a lower bound on the time averaged gas density of
the CNM.  Our model therefore cannot account for AGN.

@@

On the other hand, \citealt{LauerFaber+:2005a} find that $\sim 60\%$
of cusp galaxies and $\sim29\%$ of core galaxies have (generally
unresolved) emission in excess of the inward extrapolation of the
Nuker law.  They note that these nuclei are bluer than the background
galaxy starlight, suggesting the presence of younger stars towards
galactic center. Nuclear star cluster could be important for small
values of $M_{\rm bulge}\lsim 10^{11} \Msun$ ($\Mbh\lsim 2\times 10^8
\Msun$), but are extremely subdominant for higher mass galaxies
\citep{GrahamSpitler:2009a}.

Mass injection to the nucleus could (for $\Mbh\lsim 2 \times 10^{8}
\Msun$) therefore be dominated by a nuclear star cluster with its own
break radius on parsec scales, i.e. much smaller than our fiducial
value of $r_b \sim 100$ pc.  In the case of high heating, a smaller
break radius does not appreciably change the stagnation radius,
resulting in a similar SMBH accretion rate.  For break radii smaller
than the nominal stagnation radius, the presence of a smaller break
radius would likely decrease the stagnation radius and accretion rate.
